---
title: "02_viz"
author: "Lauren Littig"
date: "2026-10-06"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)
library(patchwork)

data("weather_df")
```

Now we have everythign we need!

```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  labs(
    title = "Temperature (Max vs Min)",
    x = "Max Tempoerature (C)",
    y = "Min Temperature (C)",
    color = "Location",
    caption = "Data from NOAAS for three weather stations."
  )
```

Now lets try some other scales.

```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  labs(
    title = "Temperature (Max vs Min)",
    x = "Max Tempoerature (C)",
    y = "Min Temperature (C)",
    color = "Location",
    caption = "Data from NOAAS for three weather stations."
  ) +
  scale_x_continuous(
    breaks = c(-10, 0, 15),
    labels = c("-10 C", "0", "Fifteen")
  ) +
  scale_y_continuous(
    trans = "sqrt",
    position = "right"
  )
```

Let's look at color!!!

```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  scale_color_hue(h = c(100, 300))
```

Color-blind friendly:

```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  viridis::scale_color_viridis(
    discrete = TRUE
  )
```

## Themes

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  ) +
  theme(legend.position = "bottom")
```

Cool! Moves legend to bottom. adding theme_classic() is cool.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  ) +
  theme_classic() +
  theme(legend.position = "bottom")
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  ) +
  theme_minimal()
```


Update the tmax vs date plot. Mine:

```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.7) +
  theme_classic() +
  theme(legend.position = "bottom")
```

His:

```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(
    title = "Seasonal trends in Max Temp",
    x = "Date",
    y = "Max Temp",
    caption = "Max daily temp in three weather stations in 2021 and 2022",
    color = "Location"
  ) +
  viridis::scale_color_viridis(
    discrete = TRUE
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Two more weird but useful plot things:

```{r}
central_park_df = 
  weather_df |> 
  filter(name == "CentralPark_NY")

molokai_df = 
  weather_df |> 
  filter(name == "Molokai_HI")

ggplot(molokai_df, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park_df)
```

multiple panels with different plot types.. COOL!!!!!

```{r}
ggp_tmax_tmin =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  theme(legend.position = "none")

ggp_prcp_density =
  weather_df |> 
  filter(prcp > 0) |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5) +
  theme(legend.position = "none")

ggp_seasonal =
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  theme(legend.position = "bottom")

(ggp_tmax_tmin + ggp_prcp_density) / ggp_seasonal
```

## Data manipulation

Start with facotrs.

boxplots!

```{r}
weather_df |> 
  mutate(name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

Reoder name variable according to another variable:
Putting name in order of median tmax. I like it!! 

```{r}
weather_df |> 
  mutate(name = fct_reorder(name, tmax)) |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

Make that distribution plot

```{r}
weather_df |> 
  select(name, tmax, tmin) |> 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temp"
  ) |> 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = 0.5) +
  facet_grid(.~name)
```

Make an FAS plot. 

```{r}
pups_df =
  read_csv(
    "data/FAS_pups.csv", skip = 3, na = c("", ".", "NA")) |> 
  janitor::clean_names()

litters_df =
  read_csv(
    "data/FAS_litters.csv", na = c("", ".", "NA")) |> 
  janitor::clean_names() |> 
  separate(group, into = c("dose", "day_of_tx"), 3)

fas_df =
  left_join(pups_df, litters_df, by = "litter_number")

fas_df |> 
  select(dose, day_of_tx, starts_with("pd")) |> 
  pivot_longer(
    starts_with("pd"),
    names_to = "outcome",
    values_to = "pn_day"
  ) |> 
  drop_na() |> 
  ggplot(aes(x = dose, y = pn_day)) +
  geom_boxplot() +
  facet_grid(day_of_tx ~ outcome)

```



